Interest Flooding Attack and Countermeasures in Named Data Networking

### 2013-03-25 paper.tex
Final pass
git-svn-id: https://ndnpub.parc.com/ndn-papers/NDN-SEC/dos-protection@941 df1a90cd-247b-4364-bdaf-287a704eb499
git-svn-id: https://ndnpub.parc.com/ndn-papers/NDN-SEC/dos-protection@944 df1a90cd-247b-4364-bdaf-287a704eb499

% \documentclass[10pt, conference, compsocconf]{IEEEtran}
\documentclass[conference]{IEEEtran}
% http://networking2013.poly.edu/call-for-papers/paper-submission/
% Page limit 9 pages

@@

@@

%\title{Mitigating Interest Flooding DDoS Attacks in NDN}

\title{Interest Flooding Attack and Countermeasures in Named Data Networking}%:\\ Cache/Content Poisoning}

%I don't think submission is anonymous...
% I PUT AUTHOR NAMES IN ALPHEBETICAL ORDER. It is the common practice in security and I always list authors alphebetically in my publications. -Ersin 
\author{
    \IEEEauthorblockN{Alexander Afanasyev\IEEEauthorrefmark{1}, Priya Mahadevan\IEEEauthorrefmark{2}, Ilya Moiseenko\IEEEauthorrefmark{1}, Ersin Uzun\IEEEauthorrefmark{2}, Lixia Zhang\IEEEauthorrefmark{1}}
    \IEEEauthorblockA{\IEEEauthorrefmark{1}University of California, Los Angeles
    \\\{afanasev, iliamo, lixia\}@cs.ucla.edu}
    \IEEEauthorblockA{\IEEEauthorrefmark{2}Palo Alto Research Center
    \\\{ersin.uzun, priya.mahadevan\}@parc.com}
}

@@

\newcommand{\todo}[1]{\vspace{2 mm}\par \noindent \marginpar{\textsc{ToDo}}
\framebox{\begin{minipage}[c]{0.95 \columnwidth}
\tt #1 \end{minipage}}\vspace{5 mm}\par}

\usepackage{graphicx}
% \usepackage[colorlinks]{hyperref}
\usepackage[]{hyperref}
\usepackage{breakurl}
\usepackage{url}
\usepackage[nocompress]{cite}
% \usepackage{verbatim}
% \usepackage{algpseudocode}
\usepackage{algpseudocode,algorithm}
% More customizeable version. Probably it would be better to convert pseudocode to 2e format
% \usepackage{algorithm2e} 
\usepackage{multirow}
\usepackage{times}
\usepackage{color}

\graphicspath{{figures/}}

@@

@@

\input{01-intro}

\input{02-ndn-background}

\input{03-interest-flooding}

\input{04-design}

\input{05-evaluation}

\input{06-related}

\input{08-conclusion}

>>>>

\documentclass[10pt,conference]{IEEEtran}
% http://networking2013.poly.edu/call-for-papers/paper-submission/
% Page limit 9 pages

@@

\hyphenation{incoming}

@@

\usepackage[top=.75in, bottom=1in, left=.625in, right=.625in]{geometry}
\usepackage{graphicx}
\usepackage[]{hyperref}
\usepackage{breakurl}
\usepackage{url}
\usepackage[nocompress]{cite}
% \usepackage{verbatim}
% \usepackage{algpseudocode}
\usepackage{algpseudocode,algorithm}
% More customizeable version. Probably it would be better to convert pseudocode to 2e format
% \usepackage{algorithm2e} 
\usepackage{multirow}
\usepackage{times}
\usepackage{color}

@@

\newcommand{\todo}[1]{\vspace{2 mm}\par \noindent \marginpar{\textsc{ToDo}}
\framebox{\begin{minipage}[c]{0.95 \columnwidth}
\tt #1 \end{minipage}}\vspace{5 mm}\par}
\graphicspath{{figures/}}

@@

\title{Interest Flooding Attack and Countermeasures in Named Data Networking}

% I PUT AUTHOR NAMES IN ALPHEBETICAL ORDER. It is the common practice in security and I always list authors alphebetically in my publications. -Ersin 
\author{
    \IEEEauthorblockN{Alexander Afanasyev\IEEEauthorrefmark{1}, Priya Mahadevan\IEEEauthorrefmark{2}, Ilya Moiseenko\IEEEauthorrefmark{1}, Ersin Uzun\IEEEauthorrefmark{2}, Lixia Zhang\IEEEauthorrefmark{1}}
    \IEEEauthorblockA{\IEEEauthorrefmark{1}University of California, Los Angeles
    \\\{afanasev, iliamo, lixia\}@cs.ucla.edu}
    \IEEEauthorblockA{\IEEEauthorrefmark{2}Palo Alto Research Center
    \\\{ersin.uzun, priya.mahadevan\}@parc.com}
}

@@

\section{Introduction}
\label{sec:intro}

Named Data Networking (NDN)~\cite{ndn-conext, ndn-tr} is an ongoing research effort that aims to move the Internet into the future with a content-centric design that is capable of efficient content distribution and seamless mobility support. In contrast to today's Internet, a key goal of the NDN project is ``security by design." In fact, it goes a long way by guaranteeing the integrity and provenance of every Data packet with digital signatures and protecting user-privacy with no source addresses carried in packets. However, one big question that is yet to be answered is: how does the NDN architecture fare in terms of its resilience against distributed denial of service (DDoS) attacks? Especially since various forms of DDoS attacks pose a significant threat to the existing Internet infrastructure~\cite{arbor-report}, it is crucial to ensure that the new design is free of similar vulnerabilities.

NDN eliminates host-based addressing and makes data the first-class network entity. Instead of sending packets to a given IP address, NDN nodes request desired data by sending Interest packets carrying application-level data names, and the network returns the requested Data packets following the path of Interests. 
Such a shift automatically eliminates several long-standing DDoS attacks, including direct flooding and reflector attacks through source address spoofing~\cite{mirkovic2004taxonomy}.
%
However, malicious users can attack the network by sending an excessive number of Interests. 
Because each Interest consumes resources at intermediate routers as it is routed through the network, an excessive number of Interests can congest the network and exhaust routers' memory resources. 
We coin the term {\it Interest flooding} to refer to such attack and
this paper exclusively investigates the problem and the solution space for it.

Our effort is an important first step towards a complete investigation of DDoS attacks in NDN. We experiment with three algorithms that allow routers to exploit their state information to thwart these attacks. Through extensive simulations, we show how one of our mitigation methodologies effectively shuts down malicious users while preventing legitimate users from service degradation. 
The rest of the paper is organized as follows. We give an overview of NDN architecture in Section~\ref{sec:ccn-intro} and explain interest flooding attacks in Section~\ref{sec:interest-flooding}. In Sections~\ref{sec:design} and \ref{sec:evaluation} we introduce techniques to mitigate these attacks, evaluate their effectiveness, and discuss their limitations. We summarize related work in Section~\ref{sec:related-work} and finalize the paper with conclusions and future work in Section~\ref{sec:conclusion}.

\section{NDN Overview\label{sec:ccn-intro}}

In this section we briefly introduce NDN with a focus on its stateful forwarding plane (refer to \cite{ndn-conext, ndn-tr, adaptive-forwarding} for more details).
NDN is a receiver-driven, data-centric communication protocol.
All communication in NDN is performed using two distinct types of packets: \textit{Interest} and \textit{Data}. Both types of packets carry a \textit{name}, which uniquely identifies a piece of data that can be carried in one Data packet. Data names in NDN are hierarchically structured and an example name for the first segment of a youtube video would look like: \ndnName{/youtube/videos/0F8YdlkKO9A/0}.

To retrieve data, a consumer requests it by sending an Interest packet with the name of the desired content in it.
Routers use this name to route the Interest towards data sources, and a Data packet whose name matches the name in the Interest is returned to the consumer by following the reverse path of the Interest. Similar to IP, Interest forwarding is based on longest name prefix match, but, unlike IP, an Interest packet and its matching Data packet always take symmetric paths.
%% An Interest is ``satisfied'' when it is matched with a Data packet.

Each NDN router maintains three major data structures:
\begin{itemize}
\item \textit{Pending Interest Table (PIT)} holds all ``not yet satisfied'' Interests that have been sent upstream towards potential data sources. Each PIT entry contains one or multiple incoming and outgoing physical interfaces; multiple incoming interfaces indicate the same data is requested from multiple downstream users; multiple outgoing interfaces indicates the same Interest is forwarded along multiple paths.
\item \textit{Forwarding Interest Base (FIB)} maps name prefixes to one or multiple physical network interfaces, defining allowed directions to where Interests can be forwarded. 
% Having one-to-many relationship in this table allows multipath forwarding of Interests.}
\item \textit{Content Store (CS)} temporarily buffers Data packets that pass through this node, allowing efficient data retrieval by different consumers.
\end{itemize}

When a router receives an Interest packet, it first checks whether there is a matching data in its CS.
If a match is found, the Data packet is sent back to the incoming interface of the Interest packet.
If not, the Interest name is checked against the entries in the PIT. 
If the name already exists in the PIT, then it can be a duplicate Interest (identified by a random number each Interest carries) that should be dropped,
or an Interest from another consumer asking for the same Data, which requires the incoming interface of this Interest to be added to the existing PIT entry.
If the name does not exist in the PIT, the Interest is added into the PIT and forwarded to the interface chosen by the strategy module, which uses FIB as input for its decisions.

When a Data packet is received, its name is used to look up the PIT.
If a matching PIT entry is found,
the router sends the Data packet to the interface(s) from which the Interest was received, caches the data in the CS, and removes the PIT entry.  Otherwise, the Data packet is deemed unsolicited and is discarded. 
Each Interest also has an associated lifetime; the PIT entry is removed when the lifetime expires.
Although the maximum lifetime is specified by users, it is ultimately a router's decision on how long it is willing to keep a PIT entry.  
For simplicity, we assume routers keep Interests in PIT for one second in our simulations.

\section{Interest Flooding Attacks in NDN}
\label{sec:interest-flooding}

As we explained earlier, Interest packets in NDN are routed through the network based on content name prefixes and consume memory resources at intermediate routers. This makes them a potential tool to launch DDoS attacks in NDN. An attacker or a set of distributed attackers can inject excessive amounts of Interests in an attempt to overload the network and cause service disruptions for legitimate users (Fig.~\ref{fig:flooding example}).

\begin{figure}[htbp]
  \centering
  \includegraphics[scale=0.39]{attack-definition}
  \vspace{-0.4cm}
  \caption{Example of the Interest flooding attack}
  \label{fig:flooding example}
\end{figure}

Since an NDN network fetches data by its name, an adversary cannot easily target specific routers or end-hosts.
However, an adversary can target a specific namespace.
For example in Fig.~\ref{fig:flooding example}, if the data producer is the exclusive owner of \ndnName{/foo/bar} namespace, both router B and the data producer would receive all Interests for \ndnName{/foo/bar/\ldots} that cannot be otherwise satisfied from in-network caches.
\footnote {This example assumes that the adversary floods the network with unique data names carrying \ndnName{\small /foo/bar} prefix to make them effective. It also assumes the producer is single-homed and the data is not replicated elsewhere. With multi-homed producers or replicated data, NDN would likely to cope better with DDoS attacks due its native multipath and adaptive forwarding~\cite{adaptive-forwarding} support.}
A large volume of such malicious Interests can disrupt service quality in NDN network in two ways: \emph{create network congestion} and \emph{exhaust resources on routers}.

Similar to packets in traditional networks, Interest packets in NDN consume a portion of network capacity. A large number of Interest packets might cause congestion and lead to legitimate packets being dropped in the network. In particular, a coordinated DDoS attack could target one specific namespace and concentrate attack traffic in certain segments of the network as routing in NDN is based on name prefixes.

As NDN routers maintain per-packet states for each forwarded Interest (i.e., an entry in its PIT), an excessive amount of malicious Interests can lead to exhaustion of a router's memory, making the router unable to create new PIT entries for incoming Interests and disrupting service for legitimate users.

Nevertheless, creating an effective Interest flooding attack in NDN is non-trivial.
To efficaciously target a specific namespace (e.g., \ndnName{/newyorktimes/}), an adversary needs to make sure that (1) the expressed Interests are routed towards and as close to the data producer/provider as possible, and (2) new corresponding PIT entries are created for those interests and are stored at intermediate NDN routers for as long as possible. The former is achieved when interests share the same name prefix (e.g., \ndnName{/newyorktimes/}) and as long as they are not served from caches of intermediate routers --an interest is not forwarded upstream if a router can satisfy it from its content store. The latter requires every single malicious Interest to ask for a unique content --all Interests requesting the same content are collapsed to one PIT entry in routers. Thus, an adversary has to request either an unpopular (i.e., not cached in routers) or non-existing unique content at each interest. Among the two options of an adversary, the first one is quite problematic due to the difficulties around indexing content names in a particular namespace, coordinating large number of bots to send unique interests and sustaining the attack while the network is continuously caching the requested content objects. However, the second option, requesting unique non-existing content at each interest, is easy to achieve and sustain. An adversary can construct such interests by concatenating a random name component to the victim namespace (e.g.,\ndnName{/newyorktimes/fgb...3rf3}) and can increase the memory usage at each router by lengthening the random component. In this paper, we exclusively focus on this particular attack strategy as it not only maximizes the damage from each malicious Interest but also is the one that is easy to launch and widely applicable to all namespaces (small or large). % other less effective strains of Interest flooding attacks can also be mitigated by applying the same or similar countermeasures described in the next section.

In the rest of this paper, we use the general term \emph{Interest flooding attack} to refer to the above described attack and assume an attacker that is limited to controlling a botnet of end-hosts only, i.e., we assume the routers in the network and the computers in the victim domain are not compromised.

%\paragraph{Assumptions} 
%E: These are assumptions for the simulations on hand and particulary to test for the worst case scenario in many aspects. They are not the paper's assumption and in fact the paper first should be more general on describing all possibilities. Then it should explain the assumptions for the simulations and discuss why they make sense and do NOT favor positive results in some way.

% - assumptions about attacker position
% - assumptions about the producer / producer namespace
% - assumptions about the attack traffic / attack pattern

%In this paper we are making the following assumptions about the Interest flooding attack:
%\begin{itemize}
%\item only client nodes can be compromised and become attack bots;
%\item there are no colluding attackers inside the Data producer's network;
%\item the attack is carried only using unique junk Interests; and
%\item there is only one Data producer for the attacked prefix.
%\end{itemize}

%We also assume that NDN forwarding strategy uses only single-path Interest forwarding, always choosing the best-metric route advertised by the routing protocol.
%This way, we are able to analyze the attack in its best environment, as enabling the multi-path forwarding would only alleviate effects of the Interest flooding attack.

%In section~\ref{sec:discussion} we discuss potential of the Interest flooding attack under several other attack assumptions.

\section{Interest flooding mitigation methods}
\label{sec:design}

% Points what should be here:
% - What can be done (in general) to mitigate flooding attack
% - Which building blocks NDN architecture gives to mitigate Interest flooding attacks: Interest limits, ability to measure Interest satisfaction performance (Interest satisfaction stats)
% - Methods to set these limits: static/dynamic
% - Methods how limits can be applied: best-effort (fifo), "fair" queuing, probabilistic
% - Reference to caching: we don't consider it here, but caching provides additional level of protection, especially for certain types of attacks

% Our definition of attack mitigation is that good clients are still able to access data on the producer.

In this section we present several algorithms to mitigate Interest flooding attacks in NDN.  Our mitigation strategies feature varying degrees of implementation complexity and effectiveness---the higher the implementation complexity, the more effective is the algorithm against Interest flooding attacks. We start by describing our simple strategies and use the insights and lessons learned from the deployment of these to inform and design more effective mitigation techniques that work well in various topologies.

%   - Naive approach to Interest Limits (that is called physical limits everywhere else)
%     Example how this can be implemented in a simple way
%     Baseline solution
%   - Interest limits with "fair" queuing
%     Improving problem of simple limits (no single face can dominate), but doesn't solve the proble

% - Attack mitigation
%   - Per-incoming interface Interest statistics

%   - Dynamic Interests limit adjustments

%   - Probabilistic Interest accept

%The fact that each Data packet takes the reverse path of the corresponding Interest packet allows intermediate routers to match a received Data packet to the corresponding Interest packet and %thus determine which Interest packets were satisfied. As we describe in the rest of this section,  NDN routers can exploit this information to make informed decisions on which Interest packets to %forward, how many Interest packets to forward, and thus effectively defend against DDoS attacks.

%Our methods to mitigate Interest flooding attack rely on the fundamental principle of the NDN architecture: the flow balance between Interest and Data packets: one Interest packet (the only communication initiator) can be satisfied with at most one Data packet.
%Because NDN is host-to-host architecture (as opposed to end-to-end in the current IP Internet), the flow balance principle allow any entity on the network, end-hosts and routers, control what and how much Data they want to receive.
%Therefore, any node can limit the number of forwarded Interests, effectively limiting the amount of the retrieved Data.
%At the same time, each forwarded Interest can be used to build up various data plane performance statistics, such as per-incoming interface ratios of satisfied Interests.

%\subsection{Na\"{i}ve attack mitigation}

%We start with a couple of naive strategies that can be used to migitate attacks in an NDN network and describe the pitfalls associated with these strategies.

% In a normal network operation, the size of Interests packets is supposed to be significantly smaller than the size of the requested Data.

% All communication in NDN network is host-to-host and receiver-driven.

An obvious and na\"ive solution to defend against Interest flooding attacks is to restrict the number of Interests that are forwarded in the network. To this end, we exploit a 
 fundamental principle of NDN architecture---flow balance between Interest and Data packets. The flow balance refers to the fact that one Interest can be satisfied by at most one Data packet. This principle allows intermediate routers to control the inbound data traffic by controlling the number of outstanding Interests in the network. 
One simple implementation technique  is for an NDN router to limit the number of forwarded Interests out of each interface based on the physical capacity of the corresponding interface. This technique is a slight modification of the well-known {\it Token Bucket} algorithm that is currently widely used in packet-switched networks. Analogous to the {\it Token Bucket} algorithm, NDN routers keep track of the amount of data requested that can fully utilize the downstream link (estimated from the number of forwarded Interests) and once the link capacity limit has been reached, they no longer forward new incoming Interests. Ideally, the number of tokens (the pending \emph{Interests Limit}) for each link will be proportional to the link's bandwidth-delay product (BDP)~\cite{tcp-survey}. We can formalize this value as follows:
% With the objective to request as many Data packets, as downstream link can pump through, we are getting the following equation for Interest limit:
%
\begin{equation}
\label{eq:interest limit}
\small \mathrm{Interest\ Limit} = Delay\ [s] \cdot \frac{\mathrm{Bandwidth\ [Bytes/s]}}{\mathrm{Data\ packet\ size\ [Bytes]}}
\end{equation}

In the above equation, \emph{Delay} is the expected time of Interest being satisfied and \emph{Data packet size} is the size of the returned Data packet.
Although both these values are not known a priori, it is not really necessary to use their exact values.
One can simply set the pending Interest limit based on the average values of round trip time and observed Data packet size, as the network buffers can smooth out most of the network fluctuations.

This {\it Token Bucket} approach might be exceptionally restrictive in forwarding Interests---not all Interests will result in a Data packet---and might result in underutilization of the network. However, the biggest drawback of this algorithm is the fact that it can nourish DDoS attacks. If a router has utilized all its tokens to forward malicious Interests, it can no longer forward incoming Interests from legitimate users till the pending malicious Interests start to expire. One way to get around the issue is to impose a per interface limit, so that malicious Interests are not allowed to entirely consume the limits of a specific interface. We describe this technique in greater detail below.

%That is, if it is known that the amount of already requested data ($=$~amount of forwarded Interests) can fully utilize the downstream link, an NDN node---either an end-user or an NDN router---has absolutely no point in forwarding new incoming Interests and creating corresponding PIT entries.
%For example, if a router A on Fig.~\ref{fig:flooding example} already forwarded 125 Interests requesting 1000-byte Data packets, any new incoming Interests can be almost safely dropped, provided the link capacity between A--B is 100~Mbps and delay is 10~ms.
%In other words, 125 Data packets returned by router B will fully utilize the link ($125 \times 1000 \mathrm{~bytes} \approx 10\mathrm{~ms} \times 100\mathrm{~Mbps}$) and any excess Data would be dropped.
%
%\subsubsection{\textbf{Interest limits (physical limits)}}
%\label{sec:physical limits}
%
%\input{algos/physical-limits.tex}

%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%

\subsubsection{\textbf{Token bucket with per interface fairness}}
\label{sec:queuing}

To address the lack of fairness associated with the na\"ive {\it Token Bucket} approach, we modify it to ensure that the Interests forwarded by a router on each interface represent a fair mix of Interests received from neighboring nodes. For example, in Fig.~\ref{fig:flooding example} router A can ensure that the tokens associated with Interests sent out on interface {\texttt eth2}  are fairly distributed across incoming interfaces \texttt{eth0} and \texttt{eth1}. 
%Due to the expected small volume of Interests, we cannot merely rely on network buffers to perform statistical multiplexing of Interests, as they would almost never be %buffered.
%At the same time, until bag of tokens is not empty, there is no reason to delay Interest forwarding, as we do not known how many and from which interfaces Interests %will arrive in the future.
In order to achieve our goal of ensuring ``fair'' mixing of Interests from all neighboring nodes, 
%, we implement an additional features to support buffering of Interests, if they cannot be immediately forwarded. 
we extend the Pending Interest Table to support flagging of Interests that cannot be immediately forwarded and implement hierarchical queues for each interface (see Fig.~\ref{fig:queueing}). 
This mechanism is essentially a class based queuing~\cite{floyd1995link}, with classes for each outgoing and incoming interface.
%As an alternative to hierarchical queues, one can also use an approach based on virtual time. For details, please see ~\cite{zhang1990virtual}.
We note that unlike normal queuing, Interest queues do not actually store a packet, but merely a bi-directional pointer to the existing PIT entry.
Thus, a PIT entry can be quickly updated when the Interest is actually forwarded, and the element can be easily removed from the queue when the Interest expires.

%For the buffering part, we can reuse Pending Interest Table, with a small extension to support flagging of the Interests that cannot be forwarded immediately (see example on Fig.~\ref{fig:queueing}). 
%As for the mixing part, we need an additional fair queuing mechanism, which can be implemented in a form of hierarchical queues (on Fig.~\ref{fig:queueing})\footnote{This essentially is a class based queuing, with classes for each outgoing/incoming interface.} or using virtual time approach~\cite{zhang1990virtual}.

\begin{figure}[thb]
  \centering
  \includegraphics[scale=0.65]{queue}
  \vspace{-0.7cm}
  \caption{Interest queuing: if tokens are unavailable, the router creates a PIT entry, but instead of forwarding, it enqueues the Interest}
  \label{fig:queueing}
\end{figure}

We present a formalized description of this algorithm in Pseudocode~\ref{alg:queuing}. 
By setting appropriate queue sizes, we can control the amount of physical resources utilized at a router.
 % (such as memory and computation power). 
It is also important to set a sensible value for how long an Interest can be enqueued. 
If an Interest is enqueued for a long time, by the time it is dequeued and forwarded, the retrieved Data packet might be dropped at the downstream routers if their corresponding state expired. 
For our evaluations, we empirically chose to enqueue Interests up to 10\% of their original lifetime.
% We believe that implementing additional mechanisms on each pair of communicating routers to keep the Interests from getting stale would mitigate this issue, though the details are beyond the scope of this paper.
% Alex: I don't believe there is a need for separate discussion about "freshness", as we already introduced concept of lifetime
% {\color{red} \it Priya:  We need to  touch upon the Interest freshness issue in the NDN overview section; else we should get rid of this bit.}

%It should be noted that enqueued Interests should not be kept in the queue for a prolonged period of time.
%Otherwise, by the time the Interests reaches the Data, the state could have been long expired downstream, effectively making such an Interest useless.
%Additional mechanisms of pair-wise agreements between NDN routers and periodic Interest refresh can solve this particular problem, but it is out of the scope of the present paper.

%The algorithm extends the base Physical limits algorithm by enabling queuing when the bag of tokens is empty (lines 7--10), as well as by triggering an action (lines 16--21), when a token becomes available and enqueued Interest can be finally forwarded.
%At the same time, the algorithm limits number of Interests allowed in a queue, constraining memory usage increase by at most a constant factor, compared to the base Physical limits algorithm (i.e., memory attack on routers are still unfeasible).

\floatname{algorithm}{\small Pseudocode}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{algorithm}[h]
\footnotesize
\caption{\small Token bucket with per interface fairness}
\label{alg:queuing}
\begin{algorithmic}[1]
\For{\textbf{each} interface \textbf{if}}
    \State{$L_{if} \leftarrow$ Interest Limit according to (1)}
    \State{$O_{if} \leftarrow 0$} \Comment{Outstanding Interests on interface \textbf{if}}
\EndFor

\vspace{0.1cm}

\Function{OutInterest}{Interest \textbf{i}, InInterface \textbf{in}, OutInterface \textbf{out}}
    \If{$L_{out} - O_{out} > 0$} \Comment{\textbf{out} is under limit cap}
        \State $O_{out} \leftarrow O_{out} + 1$  \Comment{``borrow'' a token from the bucket}
        \State add \textbf{out} to PIT entry and forward \textbf{i} to \textbf{out}
    \Else
        \State Queue $q \leftarrow out$.GetSubQueue($in$)
        \If{$Size(q) < L_{out}$}
           \State $q$.PushInterest($i$)
           \State add \textbf{out} to PIT entry, and link PIT entry with the queue
        \Else
           \State drop Interest
        \EndIf
    \EndIf
\EndFunction

\vspace{0.1cm}

\State{} \Comment{\textit{Whenever $L_{out} - O_{out}$ becomes larger than zero}}
\Function{TokenBecomesAvailable}{}
    \State Queue $q \leftarrow$ $out$.GetRoundRobinSubQueue 
    \State Interest $i \leftarrow$ $q$.PopInterest
    \State update PIT entry and Forward($i$, $out$)
\EndFunction

\vspace{0.1cm}

\Function{InData}{Data \textbf{d}}
   \State lookup PIT entry \textbf{p} for data \textbf{d}
   \For{\textbf{each} outgoing interface \textbf{out} in \textbf{p}}
        \State $O_{out} \leftarrow O_{out} - 1$ \Comment{``return'' token}
   \EndFor
\EndFunction

\vspace{0.1cm}

\Function{Timeout}{PIT entry $p$}
   \For{\textbf{each} outgoing interface \textbf{out} in \textbf{p}}
        \State $O_{out} \leftarrow O_{out} - 1$ \Comment{``return'' token}
   \EndFor
\EndFunction

\end{algorithmic}
\end{algorithm}

As we show in Section~\ref{sec:evaluation}, this algorithm provides partial relief from the Interest flooding attacks, allowing legitimate users to successfully fetch Data for 15--20\% of their expressed Interests. 
We note that while this algorithm might be reasonable for ensuring limited fairness in an NDN network, it is ineffective in protecting legitimate users from malicious ones. 
Attackers are able to successfully thwart access to content for legitimate users by sending a relatively modest volume of malicious Interests.

%At the same time, the Physical limits with or without fair queueing allows attackers to send a relatively small volume of Interests in order to significantly impact service for the legitimate users.
%Therefore, to successfully solve the problem, we need a fundamentally different, more intelligent approach, allowing localization of the attack traffic as close as possible to the attack origin.

%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%

The key drawback of the {\it Token bucket with per interface fairness} algorithm is that it still admits a relatively large number of Interests from malicious users. A considerable percentage of these malicious Interests are forwarded all the way to content producers, thereby reducing resources available to serve legitimate users.  This algorithm attempts to ensure that each interface does not forward more than its fair share of Interests, but in doing so, it drops both legitimate and malicious Interests. For any strategy to be effective in defending against Interest flooding attacks, it must be able to detect and differentiate malicious requests from legitimate ones. 
% `Good' Interests from legitimate users must be admited and forwarded appropriately. 
Thus, the key question is how can we devise mitigation algorithms that allow a router to distinguish between `good' and `bad' Interests?

%One can try a black- or whitelisting approach. However, besides the fact that source black- and whitelisting cannot work in %NDN as it does not feature source addresses, it requires extraneous knowledge to classify the incoming Interests.

%Priya: I have added the below point specifically to point out that a router cannot determine from the prefix name alone whether /parc is good  or bad...

%{\color{red} \textbf{Alex: This paragraph is confusing to me. I would remove it.} In NDN one cannot easily distinguish a `good' prefix from a `bad' one. Content producers must be able to support %dynamic generation of Data packets to satisfy incoming Interest packets---thus, an Interest that cannot be satisfied by matching Data from any intermediate router's cache is necessarily forwarded %to the producer for possible dynamic Data packet generation. 
%Thus intermediate routers cannot determine if the prefix is `good' or `bad' based on the prefix name alone---DDoS mitigation techniques that rely on whitelisting or blacklisting of certain %namespaces  will be inefficient and likely ineffective. }

% Alex: I'm not sure that this part is relevant... The point I see is that blacklisting and whitelisting is not 100% effective, but it can be used in both worlds

% In today's host-based Internet architecture where DDoS mitigation techniques rely on blacklisting of ``dark'' IP prefixes or whitelisting of good IP prefixes.
% Unlike IP, in NDN one cannot easily distinguish a `good' prefix from a `bad' one. 
% Content producers must be able to support dynamic generation of Data packets to satisfy incoming Interest packets---thus, an Interest that cannot be satisfied by matching Data from any intermediate router's cache is necessarily forwarded to the producer for possible dynamic Data packet generation. 
% Thus intermediate routers cannot determine if the prefix is `good' or `bad' based on the name alone---DDoS mitigation techniques that rely on whitelisting or blacklisting of certain namespaces  will be inefficient and likely ineffective.

\subsection{Intelligent attack mitigation}
\label{sec:intelligent mitigating}

%While the Interest limit is the key building block to suppress mechanisms of the Interest flooding attack, legitimate Interests need to be somehow prioritized and malicious Interests need to be somehow penalized in order to completely suppress the attack.
%That is, instead of processing Interests always based on the first-in-first-serve rule, NDN routers need some basis to treat the incoming Interests differently.
%Thus, the primary task in bringing intelligence to the Interest flooding attack mitigation is to proactively distinguish between legitimate and malicious Interests.

In order to distinguish between legitimate and malicious Interests, we leverage another unique feature of  NDN architecture---guaranteed symmetric flow of Interest and Data packets. Since a Data packet takes the reverse path of the corresponding Interest packet, a router is guaranteed to see if an Interest it forwarded resulted in a matching Data packet or timed out. %The only exception being if Interest/Data packets were lost along the way due to congestion in the network.
Since malicious Interests are not likely to bring data back (as discussed in Section~\ref{sec:interest-flooding}), this information can be utilized by routers in differentiating attack and legitimate traffic.  %Therefore, intermediate routers can classify the ones that brings data back as legitimate, while the ones that timed out can be marked as malicious.\footnote{Recall that in order to maximize effect of the Interest flooding attack, an adversary expresses a large volume of junk Interests (see Section~\ref{sec:interest-flooding}).}  
%Implications of other types of attacks are discussed in Section~\ref{sec:discussion}.}

This timeout-based differentiation method is reactive in nature: one cannot determine in advance if an Interest will result in a timeout or Data being retrieved. However, routers can proactively maintain up-to-date statistics of Interest satisfaction ratios (number of forwarded versus number of satisfied Interests), and use this statistic to determine whether an incoming Interest should be forwarded or dropped. For example, maintaining independent Interest satisfaction ratio statistics for each incoming interface is sufficient to reasonably predict whether an Interest received from a neighbor connected to this interface will result in a Data packet or a timeout if forwarded. Statistics can also be kept at finer granularities such as per outgoing interface, per name prefix, etc. that can further improve the estimates. A router's goal should be to prioritize Interests that bring Data back while quickly penalizing those that occupy resources but don't generate a Data stream in return. In order to allow negative statistics to build up fast and positive statistics deteriorate quickly, we use the standard exponentially weighted moving average, performed once a second with $\alpha$ coefficient $e^{-1/30}$, approximately corresponding to a 30-second averaging window.

%Choosing the right balance between these contradictory requirements is a challenge and we explore this topic further  in the evaluation section.  %{\color{red} \it Priya: we should address the one that works best such as exponentially weighted moving average in eval section - it doesn't belong %here.}

%The devil is always in the details.
%From the one hand, such statistics needs to start penalizing adversaries as soon as possible (i.e., negative stats should build up fast).
%On the other hand, the positive statistics should not deteriorate too fast (i.e., positive stats should be relatively long-term).
%Our preliminary experiments showed that the standard exponentially weighted moving average, performed once a second with $\alpha$ coefficient $e^{-1/30}$, approximately corresponding to a 30-second averaging window, provides a good balance between the two contradictory requirements.

% \subsubsection{\textbf{Data plane performance tracking}}
% \label{sec:stats}

% Note that there is a condition (line 6 in Pseudocode~\ref{alg:probabilistic model}) to check if there is a valid statistics point.
% This condition is extremely important, because it first provides a basis to distinguish between known facts (i.e., good or bad satisfaction ratio for the incoming interface) and unknown facts (e.g., the first time an Interests arrives on the interfaces).
% Second, it gives an opportunity to recover from a bad history (history of unsatisfied Interests) after malicious Interests are ceased to flow in.
% Essentially, this recovery relies on statistics module to perform time-based invalidation of historical data (timely, but not too quickly\footnote{Otherwise, attackers may send short bursts of malicious Interests, successfully avoiding differential Interest treatment}).

Pseudocode~\ref{algo:interest stats} formally defines how statistics can be generated for  each incoming interface. Note that in order to ensure decaying of relative statistics (e.g., ratio between the number of unsatisfied and forwarded Interest), only unsatisfied statistics needs to be exponentially smoothed (lines 23--26).

\floatname{algorithm}{\small Pseudocode}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{algorithm}[h]
\footnotesize
\caption{\small Interest satisfaction statistics}
\label{algo:interest stats}
\begin{algorithmic}[1]

\For{\textbf{each} interface \textbf{if}}
    \State $F_{if} \leftarrow 0$ \Comment{forwarded Interests from interface \textbf{if}}
    \State $\hat F_{if} \leftarrow 0$ \Comment{averaged value of $F_{if}$}

    \State $U_{if} \leftarrow 0$ \Comment{unsatisfied Interests from interface \textbf{if}}
    \State $\hat U_{if} \leftarrow 0$ \Comment{averaged value of $U_{if}$}
\EndFor

\vspace{0.1cm}
\Function{OutInterest}{Interest \textbf{i}, InInterface \textbf{in}}
  \State $F_{in} \leftarrow F_{in} + 1$
  \State record \textbf{\emph{in}} in the list of incoming interfaces for \textbf{\emph{i}}
\EndFunction

\vspace{0.1cm}
\Function{InterestTimeout}{Interest \textbf{i}}
    \State lookup the list of incoming interfaces for \textbf{\emph{i}}

    \For{\textbf{each} interface \textbf{if} in the list}
        \State $U_{if} \leftarrow U_{if} + 1$
    \EndFor
\EndFunction

\vspace{0.1cm}

\State {} \Comment{\textit{Exponentially weighted moving average smoothing}}
\Function{EWMA}{} \Comment{Every second}
\State $\alpha \leftarrow e^{-1.0/30.0}$  %\Comment{$\approx$ 30~sec average}

\For{\textbf{each} interface \textbf{if}}
    \State $\hat U_{if} \leftarrow \alpha \cdot \hat U_{if} + (1 - \alpha) \cdot U_{if}$ 
    \State $U_{if} \leftarrow 0$
